\textbf{COMPARISON OF \\ HEURISTIC ALGORITHMS \\ FOR KP01} \\[0.5cm] \rule{12cm}{0.3mm} \\[0.5cm] \small \scshape{A comparison of heuristic algorithms for solving the 0-1 Knapsack Problem} \\[0.4cm] \rule{12cm}{0.3mm} \vskip -0.3cm

### 2023-05-20 doc/paper.tex
added abstract partly, formatted layout

\newpage
\tableofcontents
\listofalgorithms
\newpage
\pagenumbering{arabic}

@@

\subsection{Background}

@@

\subsection{Aim}
The aim of this paper is to compare algorithms in order to investigate the strength of commonly used techniques in for solving optimization problems.

@@

\subsection*{Deterministic algorithms}
Dynamic programming is a general technique for solving optimization problems. If a problem has an optimal substructure and over-lapping subproblems, then dynamic programming is applicable \cite{optimalSubstructure}. In computer science a problem has \emph{optimal substructure} if an optimal solution for a problem can be constructed from optimal solutions of its sub-problems, and \emph{over-lapping subproblems} is when a problem can be decomposed into sub-problems which are reused. Dynamic programming breaks down a complicated problem into smaller sub-problems in a recursive manner, while also using some memory to save the solutions of the sub-problems (usually in tabular form). This way, when one needs to get a solution for a sub-problem again, one can just use previously calculated values.

@@

3 testgroups of 5 testcases were made using a random number generator, where a random number was generated using the Mersenne Prime Twister (mt19937). A computer cannot generate a truly random number. A pseudorandom number is what is generated, which is a number which appears to be statistically random, but has been generated using a deterministic process. The 3 groups had item counts of 100, 1000, and 10000. The number of solutions in memory (for the heuristic algorithms) was put in the input file which was kept constant, 20, in all testcases and algorithms for this report. In the random number generated testcases, the number of items $n$ was chosen to either be 100, 1000 or 10000. The profit and weight values were then randomly generated in a range of $ [1, noItems] $ while a sum-of-weights variable \emph{sum} was kept. The max weight was then randomly generated in a range of $[\frac{\text{sum}}{100} , \frac{\text{sum}}{10}]$, so that the knapsack wouldn't be able to hold all the items. To avoid any overflow errors, all generated numbers (and the sum of the profit/weight values) were kept below the max value of an \emph{int} which is $2^{31}-1$ in many programming languages.  Then $n$, $W$ (maxWeight), and each item (profit and weight values) was put in an input file. To make a fair comparison the number of $ITERATIONS$ (how many times a candidate solution is generated) was set to 100 for both algorithms.

@@

\subsection{Raw Data}
Refer to the appendix for the \hyperlink{random100}{raw data} for the algorithms running on the six testgroups: 100, 1000, 10k randomized numbers respectively and 3 differently correlated datasets. 
% \href{https://github.com/raunakCode/diploma-project/blob/main/testing/algorithms%20results.csv}{\textcolor{purple}{Link to results}}

@@

\begin{changemargin}{8mm}{8mm} 
\begin{table}[h!]
    \centering
    \scriptsize
    \caption{Mean and StDev. of 100 Randomized items} \label{100-mean}
    \begin{tabu}{|c|[1pt]c|c|c|c|c|c|}
        \tabucline[1pt]{1-6} 
        Testcase & 1 & 2 & 3 & 4 & 5 \\ [-1pt] \tabucline[1pt]{1-6} 
        DGHS Mean &4160.33$\pm$3.50 &2561$\pm$0 &5288$\pm$19.66 &3535$\pm$0 &1747$\pm$0 \\[1em] \hline
        Mean Percent deviation from optimal &0.21\%  &0\%  &0.40\%  &0\%  &0\%  \\ [-1pt] \tabucline[1pt]{1-6}
        BMQHOA Mean &3926.83$\pm$40.27 &2561.00$\pm$0 &4879$\pm$46.18 &3360.66$\pm$68.96 &1711.5$\pm$25.71 \\[1em] \hline
        Mean Percent deviation from optimal &5.80\% &0\%  &8.09\% &4.93\% &2.03\% \\ [-1pt] \tabucline[1pt]{1-6}
    \end{tabu}

    \vspace{1cm}
    \hspace{-4cm}
    \caption{Mean and StDev. of 1000 Randomized items} \label{1000-mean}
    \begin{tabu}{|c|[1pt]c|c|c|c|c|c|}
        \tabucline[1pt]{1-6} 
        Testcase & 1 & 2 & 3 & 4 & 5 \\ [-1pt] \tabucline[1pt]{1-6} 
        DGHS Mean &156083.83$\pm$830.04 &118603.83$\pm$830.04 &144070$\pm$1343.77 &109734.33$\pm$1037.03 &126366.16$\pm$1236.73 \\ \hline
        Mean Percent deviation from optimal &13.71\% &11.76\% &14.12\% &10.77\% &13.27\% \\ [-1pt] \tabucline[1pt]{1-6}
        BMQHOA Mean &145086.5$\pm$1035.45 &109876.33$\pm$1540.63 &136622.66$\pm$2042.95 &101951.33$\pm$1271.88 &118291.16$\pm$799.78 \\ \hline
        Mean Percent deviation from optimal &19.79\% &18.25\% &18.56\% &17.10\% &18.82\% \\ [-1pt] \tabucline[1pt]{1-6}
    \end{tabu}

    \vspace{1cm}

    \caption{Mean and StDev. of 10k Randomized items} \label{10k-mean}
    \begin{tabu}{|c|[1pt]c|c|c|c|c|c|}
        \tabucline[1pt]{1-6} 
        Testcase & 1 & 2 & 3 & 4 & 5 \\ [-1pt] \tabucline[1pt]{1-6} 
        DGHS Mean &7681910$\pm$51555 &10567302$\pm$23956 &8014906$\pm$39449.15 &9347061$\pm$56967.66 &9583139$\pm$55816.60 \\ \hline
        Mean Percent deviation from optimal &22.61\% &23.11\% &22.07\% &23.10\% &23.00\% \\[-1pt] \tabucline[1pt]{1-6}
        BMQHOA Mean &7438487$\pm$29887.17 &10276681$\pm$50507.35 &7774520$\pm$34559.10 &9075476$\pm$53168.02 &9330322$\pm$46301.35 \\ \hline
        Mean Percent deviation from optimal &25.07\% &25.22\% &24.41\% &25.33\% &25.03\% \\[-1pt] \tabucline[1pt]{1-6}
    \end{tabu}
\end{table}

   % \vspace{1cm}
\begin{table}[h!]
    \centering
    \scriptsize
    \caption{Mean and StDev. of Uncorrelated Dataset} \label{uncorrelated-mean}
    \begin{tabu}{|c|[1pt]c|c|c|c|c|c|}
        \tabucline[1pt]{1-6} 
        Testcase & 1 & 2 & 3 & 4 & 5 \\ [-1pt] \tabucline[1pt]{1-6} 
        DGHS Mean &9147$\pm$0 &11238$\pm$0 &51019.50$\pm$437.85 &99036.16$\pm$1544.34 &448975$\pm$1135.46 \\ \hline
        Mean Percent deviation from optimal &0\% &0.00\% &6.39\% &10.47\% &20.34\% \\[-1pt] \tabucline[1pt]{1-6}
        BMQHOA Mean &8974.5$\pm$147.24 &10976.50$\pm$154.37 &47401.50$\pm$841.75 &91725.83$\pm$1346.23 &431259.66$\pm$3049.43 \\ \hline
        Mean Percent deviation from optimal &1.88\% &2.32\% &13.02\% &17.08\% &23.48\% \\[-1pt] \tabucline[1pt]{1-6}
    \end{tabu}

    \vspace{1cm}

    \caption{Mean and StDev. of Weakly Correlated Dataset} \label{weakly-mean}
    \begin{tabu}{|c|[1pt]c|c|c|c|c|c|}
        \tabucline[1pt]{1-6} 
        Testcase & 1 & 2 & 3 & 4 & 5 \\ [-1pt] \tabucline[1pt]{1-6} 
        DGHS Mean &1514$\pm$0 &1627.66$\pm$2.42 &8907.33$\pm$27.56 &17371.83$\pm$126.76 &82681.33$\pm$390.98 \\ \hline
        Mean Percent deviation from optimal &0\% &0.38\% &1.59\% &3.76\% &8.33\% \\[-1pt] \tabucline[1pt]{1-6}
        BMQHOA Mean &1508.16$\pm$6.82 &1613.66$\pm$16.74 &8435.33$\pm$60.60 &16539$\pm$133.95 &80900.66$\pm$970.71 \\ \hline
        Mean Percent deviation from optimal &0.38\% &1.24\% &6.81\% &8.37\% &10.31\% \\[-1pt] \tabucline[1pt]{1-6}
    \end{tabu}

    \vspace{1cm}

    \caption{Mean and StDev. of Strongly Correlated Dataset} \label{strongly-mean}
    \begin{tabu}{|c|[1pt]c|c|c|c|c|c|}
        \tabucline[1pt]{1-6} 
        Testcase & 1 & 2 & 3 & 4 & 5 \\ [-1pt] \tabucline[1pt]{1-6} 
        DGHS Mean &2391.66$\pm$6.08 &2697$\pm$0 &13682.5$\pm$60.05 &26444.66$\pm$240.84 &126448$\pm$615.61 \\ \hline
        Mean Percent deviation from optimal &0.22\% &0\% &4.91\% &8.55\% &13.93\% \\[-1pt] \tabucline[1pt]{1-6}
        BMQHOA Mean &2327.33$\pm$48.55 &2658.66$\pm$48.81 &13334.66$\pm$182.91 &25604$\pm$91.32 &123813.83$\pm$773.19 \\ \hline
        Mean Percent deviation from optimal &2.90\% &1.42\% &7.33\% &11.46\% &15.72\% \\[-1pt] \tabucline[1pt]{1-6}
    \end{tabu}
\end{table}

\begin{table}[h!]
    \centering
    \caption{Mean percentage deviation from optimal across all datasets}\label{deviation-table}
    \begin{tabu}{|c|c|c|}
        \tabucline[1pt]{1-6} 
        & DGHS mean percentage deviation & BMQHOA mean percentage deviation \\ [-1pt] \tabucline[1pt]{1-6} 
        100 randomized items &0.12\% &4.17\% \\ \hline
        1000 randomized items &12.73\%&18.50\% \\\hline
        10k randomized items &22.78\%&25.01\% \\\hline
        Uncorrelated dataset &7.44\%&11.56\% \\\hline
        Weakly Correlated dataset &2.81\%&5.42\% \\\hline
        Strongly Correlated Dataset &5.52\%&7.77\% \\[-1pt] \tabucline[1pt]{1-6} 
    \end{tabu}
\end{table}
\end{changemargin}
%100 randomized items mean and stdev graph

@@

        \legend {DGHS, BMQHOA};
        \end{axis}
    \end{tikzpicture}
    \caption{Mean percentage deviation from optimal across all datasets} \label{deviation-graph}
    \vspace{1em}
    \begin{tikzpicture}
        \begin{axis} [
            title = Means of total profit values generated by DGHS and BMQHOA for 100 randomized items,
            ybar = .05cm,
            bar width = 12pt,
            xmin = 1,
            xmax = 5,
            ymin=0,
            xlabel = Testcase,
            ylabel = Mean profit value,
            legend cell align={left},
            legend pos = north west,
            %ytick = data,
            y tick label style={
            /pgf/number format/1000 sep=},
            enlarge y limits = {value = .25, upper},
            enlarge x limits = {abs = .8}
        ]
        \addplot+ [
                error bars/.cd,
                    y dir=both,
                    % (changed from `y explicit` so the error bars are (clearly) visible
                    y explicit ,
            ] coordinates {
                (1, 4160.33) +- (0,3.5)
                (2, 2561) +- (0,0)
                (3, 5288) +- (0,19.667)
                (4,3535) +- (0,0)
                (5,1747) +- (0,0)
            };
        \addplot+ [
                error bars/.cd,
                    y dir=both,
                    % (changed from `y explicit` so the error bars are (clearly) visible
                    y explicit ,
            ] coordinates {
                (1, 3926.83) +- (0,40.27613023)
                (2, 2561) +- (0,0)
                (3, 4879) +- (0,46.18224767)
                (4,3360.66) +- (0,68.9627919)
                (5,1711.5) +- (0,25.71186497)
            };
        \legend {DGHS, BMQHOA};
        \end{axis}
    \end{tikzpicture}
    \caption{Shows the means of profit values for 100 randomized items} \label{100-graph}
\end{figure}

>>>>

\begin{abstract}
    The 0-1 Knapsack problem is a well-known NP-complete problem. Various types of heuristic algorithms like genetic algorithms, simulated annealing algorithms, population evolution algorithms have been designed to solve the 0-1 Knapsack problem. The diploma project compared two such heuristic algorithms' ability to solve the 0-1 Knapsack Problem based on the generated total profit values: the Binary Multi-Scale Harmonic Oscillator Algorithm versus the Discrete Global-Best Harmony Search Algorithm. All codes for this paper are available at \href{github.com/raunakCode}{https://github.com/raunakCode/diploma-project}.
\end{abstract}

\pagenumbering{roman}

\tableofcontents
\listofalgorithms
\newpage

@@

\pagenumbering{arabic}
\subsection{Background}

@@

\subsection{Aim}
The aim of this paper is to compare the Discrete Global-Best Harmony Search Algorithm and the Binary Harmonic Multi-Scale Algorithm on their ability to solve KP01 to investigate the strength of commonly used techniques for solving optimization problems.

@@

\subsection*{Deterministic algorithms}
Dynamic programming is a general technique for solving optimization problems. If a problem has an optimal substructure and over-lapping subproblems, then dynamic programming is applicable . In computer science a problem has \emph{optimal substructure} if an optimal solution for a problem can be constructed from optimal solutions of its sub-problems, and \emph{over-lapping subproblems} is when a problem can be decomposed into sub-problems which are reused \cite{optimalSubstructure}. Dynamic programming breaks down a complicated problem into smaller sub-problems in a recursive manner, while also using some memory to save the solutions of the sub-problems (usually in tabular form). This way, when one needs to get a solution for a sub-problem again, one can just use previously calculated values.

@@

3 testgroups of 5 testcases were made using a random number generator, where a random number was generated using the Mersenne Prime Twister (mt19937). A computer cannot generate a truly random number. A pseudorandom number is what is generated, which is a number which appears to be statistically random, but has been generated using a deterministic process. The 3 groups had item counts of 100, 1000, and 10000. The number of solutions in memory (for the heuristic algorithms) was put in the input file which was kept constant, 20, in all testcases and algorithms for this report. In the random number generated testcases, the number of items $n$ was chosen to either be 100, 1000 or 10000. The profit and weight values were then randomly generated in a range of $ [1, noItems] $ while a sum-of-weights variable \emph{sum} was kept. The max weight was then randomly generated in a range of $[\frac{\text{sum}}{100} , \frac{\text{sum}}{10}]$, so that the knapsack would not be able to hold all the items. To avoid any overflow errors, all generated numbers (and the sum of the profit/weight values) were kept below the max value of an \emph{int} which is $2^{31}-1$ in many programming languages.  Then $n$, $W$ (maxWeight), and each item (profit and weight values) was put in an input file. To make a fair comparison the number of $ITERATIONS$ (how many times a candidate solution is generated) was set to 100 for both algorithms.

@@

\subsection{Raw Data}
Refer to the appendix for the \hyperlink{random100}{raw data} for the algorithms running on the six testgroups: 100, 1000, 10k randomized numbers respectively and 3 differently profit-weight correlated datasets. 
% \href{https://github.com/raunakCode/diploma-project/blob/main/testing/algorithms%20results.csv}{\textcolor{purple}{Link to results}}

@@

\begin{changemargin}{8mm}{8mm} 
%100
\begin{table}[h!]
    \centering
    \scriptsize
    \caption{\scriptsize Shows the mean and StDev. of the total profit values produced in dataset of 100 randomized items} \label{100-mean}
    \begin{tabu}{|c|[1pt]c|c|c|c|c|c|}
        \tabucline[1pt]{1-6} 
        Testcase & 1 & 2 & 3 & 4 & 5 \\ [-1pt] \tabucline[1pt]{1-6} 
        DGHS Mean &4160.33$\pm$3.50 &2561$\pm$0 &5288$\pm$19.66 &3535$\pm$0 &1747$\pm$0 \\[1em] \hline
        Mean Percent deviation from optimal &0.21\%  &0\%  &0.40\%  &0\%  &0\%  \\ [-1pt] \tabucline[1pt]{1-6}
        BMQHOA Mean &3926.83$\pm$40.27 &2561.00$\pm$0 &4879$\pm$46.18 &3360.66$\pm$68.96 &1711.5$\pm$25.71 \\[1em] \hline
        Mean Percent deviation from optimal &5.80\% &0\%  &8.09\% &4.93\% &2.03\% \\ [-1pt] \tabucline[1pt]{1-6}
    \end{tabu}
\end{table}

\begin{figure}[hbt!]
    \vspace{1em}
    \begin{tikzpicture}
        \begin{axis} [
            title = Means of total profit values generated by DGHS and BMQHOA for 100 randomized items,
            ybar = .05cm,
            bar width = 12pt,
            xmin = 1,
            xmax = 5,
            ymin=0,
            xlabel = Testcase,
            ylabel = Mean profit value,
            legend cell align={left},
            legend pos = north west,
            %ytick = data,
            y tick label style={
            /pgf/number format/1000 sep=},
            enlarge y limits = {value = .25, upper},
            enlarge x limits = {abs = .8}
        ]
        \addplot+ [
                error bars/.cd,
                    y dir=both,
                    % (changed from `y explicit` so the error bars are (clearly) visible
                    y explicit ,
            ] coordinates {
                (1, 4160.33) +- (0,3.5)
                (2, 2561) +- (0,0)
                (3, 5288) +- (0,19.667)
                (4,3535) +- (0,0)
                (5,1747) +- (0,0)
            };
        \addplot+ [
                error bars/.cd,
                    y dir=both,
                    % (changed from `y explicit` so the error bars are (clearly) visible
                    y explicit ,
            ] coordinates {
                (1, 3926.83) +- (0,40.27613023)
                (2, 2561) +- (0,0)
                (3, 4879) +- (0,46.18224767)
                (4,3360.66) +- (0,68.9627919)
                (5,1711.5) +- (0,25.71186497)
            };
        \legend {DGHS, BMQHOA};
        \end{axis}
    \end{tikzpicture}
    \caption{Shows the means of profit values for 100 randomized items} \label{100-graph}
\end{figure}

%1000
\begin{table}
    \centering
    \scriptsize
    \caption{\scriptsize Shows the mean and StDev. of the total profit values produced in dataset of 1000 randomized items} \label{1000-mean}
    \begin{tabu}{|c|[1pt]c|c|c|c|c|c|}
        \tabucline[1pt]{1-6} 
        Testcase & 1 & 2 & 3 & 4 & 5 \\ [-1pt] \tabucline[1pt]{1-6} 
        DGHS Mean &156083.83$\pm$830.04 &118603.83$\pm$830.04 &144070$\pm$1343.77 &109734.33$\pm$1037.03 &126366.16$\pm$1236.73 \\ \hline
        Mean Percent deviation from optimal &13.71\% &11.76\% &14.12\% &10.77\% &13.27\% \\ [-1pt] \tabucline[1pt]{1-6}
        BMQHOA Mean &145086.5$\pm$1035.45 &109876.33$\pm$1540.63 &136622.66$\pm$2042.95 &101951.33$\pm$1271.88 &118291.16$\pm$799.78 \\ \hline
        Mean Percent deviation from optimal &19.79\% &18.25\% &18.56\% &17.10\% &18.82\% \\ [-1pt] \tabucline[1pt]{1-6}
    \end{tabu}
\end{table}

\begin{figure}[hbt!]
\centering
    \begin{tikzpicture}
        \begin{axis} [
            title = {Means of total profit values generated by DGHS and BMQHOA for 1000 randomized items},
            ybar = .05cm,
            bar width = 12pt,
            xmin = 1,
            xmax = 5,
            ymin = 0,
            xlabel = Testcase,
            ylabel = Mean profit value,
            legend cell align={left},
            legend pos = north west,
            %ytick = data,
            y tick label style={
            /pgf/number format/1000 sep=},
            enlarge y limits = {value = .25, upper},
            enlarge x limits = {abs = .8}
        ]
        \addplot+ [
                error bars/.cd,
                    y dir=both,
                    % (changed from `y explicit` so the error bars are (clearly) visible
                    y explicit ,
            ] coordinates {
                (1, 156083.83) +- (0,830.04)
                (2, 118603.83) +- (0,1904.70)
                (3, 144070.00) +- (0,1343.77)
                (4, 109734.33) +- (0,1037.03)
                (5, 126366.17) +- (0,1236.73)
            };
        \addplot+ [
                error bars/.cd,
                    y dir=both,
                    % (changed from `y explicit` so the error bars are (clearly) visible
                    y explicit ,
            ] coordinates {
                (1, 145086.50) +- (0,1035.46)
                (2, 109876.33) +- (0,1540.63)
                (3, 136622.67) +- (0,2042.95)
                (4,101951.33) +- (0,1271.88)
                (5,118291.17) +- (0,799.79)
            };
        \legend {DGHS, BMQHOA};
        \end{axis}
    \end{tikzpicture}
\caption{Shows the means of profit values for 1000 randomized items} \label{1000-graph}
    \vspace{1em}
\end{figure}
\vskip -5cm
%10k
\begin{table}
    \centering
    \scriptsize
    \vspace{1cm}
    \caption{\scriptsize Shows the mean and StDev. of the total profit values produced in dataset of 10k randomized items} \label{10k-mean}
    \begin{tabu}{|c|[1pt]c|c|c|c|c|c|}
        \tabucline[1pt]{1-6} 
        Testcase & 1 & 2 & 3 & 4 & 5 \\ [-1pt] \tabucline[1pt]{1-6} 
        DGHS Mean &7681910$\pm$51555 &10567302$\pm$23956 &8014906$\pm$39449.15 &9347061$\pm$56967.66 &9583139$\pm$55816.60 \\ \hline
        Mean Percent deviation from optimal &22.61\% &23.11\% &22.07\% &23.10\% &23.00\% \\[-1pt] \tabucline[1pt]{1-6}
        BMQHOA Mean &7438487$\pm$29887.17 &10276681$\pm$50507.35 &7774520$\pm$34559.10 &9075476$\pm$53168.02 &9330322$\pm$46301.35 \\ \hline
        Mean Percent deviation from optimal &25.07\% &25.22\% &24.41\% &25.33\% &25.03\% \\[-1pt] \tabucline[1pt]{1-6}
    \end{tabu}
\end{table}

\begin{figure}[hbt!]
\centering
    \begin{tikzpicture}
        \begin{axis} [
            title = Means of total profit values generated by DGHS and BMQHOA for 10k randomized items,
            ybar = .05cm,
            bar width = 12pt,
            xmin = 1,
            xmax = 5,
            ymin = 0,
            xlabel = Testcase,
            ylabel = Mean profit value,
            %ytick = data,
            legend cell align={left},
            legend pos = north west,
            y tick label style={
            /pgf/number format/1000 sep=},
            enlarge y limits = {value = .25, upper},
            enlarge x limits = {abs = .8}
        ]
        \addplot+ [
                error bars/.cd,
                    y dir=both,
                    % (changed from `y explicit` so the error bars are (clearly) visible
                    y explicit ,
            ] coordinates {
                (1, 7681910.333) +- (0, 51555.40)
                (2, 10567302) +- (0, 23956.32)
                (3, 8014906) +- (0, 39449.15)
                (4, 9347061.67) +- (0, 56967.67)
                (5, 9583139.17) +- (0, 55816.61)
            };
        \addplot+ [
                error bars/.cd,
                    y dir=both,
                    % (changed from `y explicit` so the error bars are (clearly) visible
                    y explicit ,
            ] coordinates {
                (1, 7438487.17) +- (0, 29887.18)
                (2, 10276681.83) +- (0, 50507.36)
                (3, 7774520.83) +- (0, 34559.11)
                (4, 9075476.00) +- (0, 53168.02)
                (5, 9330322.17) +- (0, 46301.35)
            };
        \legend {DGHS, BMQHOA};
        \end{axis}
    \end{tikzpicture}
\caption{Shows the mean of profit values for 10k randomized items} \label{10k-graph}
\end{figure}

%uncorrelated
\begin{table}[h!]
    \centering
    \scriptsize
    \caption{\scriptsize Shows the mean and StDev. of the total profit values produced in the profit-weight uncorrelated dataset} \label{uncorrelated-mean}
    \begin{tabu}{|c|[1pt]c|c|c|c|c|c|}
        \tabucline[1pt]{1-6} 
        Testcase & 1 & 2 & 3 & 4 & 5 \\ [-1pt] \tabucline[1pt]{1-6} 
        DGHS Mean &9147$\pm$0 &11238$\pm$0 &51019.50$\pm$437.85 &99036.16$\pm$1544.34 &448975$\pm$1135.46 \\ \hline
        Mean Percent deviation from optimal &0\% &0.00\% &6.39\% &10.47\% &20.34\% \\[-1pt] \tabucline[1pt]{1-6}
        BMQHOA Mean &8974.5$\pm$147.24 &10976.50$\pm$154.37 &47401.50$\pm$841.75 &91725.83$\pm$1346.23 &431259.66$\pm$3049.43 \\ \hline
        Mean Percent deviation from optimal &1.88\% &2.32\% &13.02\% &17.08\% &23.48\% \\[-1pt] \tabucline[1pt]{1-6}
    \end{tabu}
\end{table}

\begin{figure}[hbt!]
    \centering
    \begin{tikzpicture}
        \begin{axis} [
            title = Means of total profit values generated by DGHS and BMQHOA for the items uncorrelated by profit-weight,
            ybar,
            bar width = 12pt,
            width = 14cm,
            height = 6cm,
            xlabel = Testcase,
            ylabel = Mean profit value,
            xmin = 1,
            xmax = 5,
            ymin = 0,
            legend cell align={left},
            legend pos = north west,
            %x tick label style={rotate=45},
            xtick = data,
            ymin=0,
            %ytick = data,
            y tick label style={/pgf/number format/1000 sep=},
            enlarge y limits = {value = .25, upper},
            enlarge x limits = {abs = .8}
            %enlarge x limits = {abs = .8}
        ]
        \addplot+ [
                error bars/.cd,
                    y dir=both,
                    % (changed from `y explicit` so the error bars are (clearly) visible
                    y explicit ,
            ] coordinates {
                (1, 9147) +- (0, 0.00)
                (2, 11238.00) +- (0, 0.00)
                (3, 51019.50) +- (0, 437.85)
                (4, 99036.17) +- (0, 1544.34)
                (5, 448975.00) +- (0, 1135.46)
            };
        \addplot+ [
                error bars/.cd,
                    y dir=both,
                    % (changed from `y explicit` so the error bars are (clearly) visible
                    y explicit ,
            ] coordinates {
                (1, 8974.50) +- (0, 147.24)
                (2, 10976.50) +- (0, 154.37)
                (3, 47401.50) +- (0, 841.76)
                (4, 91725.83) +- (0, 1346.24)
                (5, 431259.67) +- (0, 3049.43)
            };
        \legend {DGHS, BMQHOA};
        \end{axis}
    \end{tikzpicture}
\caption{Shows the means of profit values for uncorrelated items} \label{uncorrelated-graph}
\end{figure}
\newpage
%weakly-correlated
\begin{table}[h!]
    \centering
    \scriptsize
    \caption{\scriptsize Shows the mean and StDev. of the total profit values produced in the profit-weight weakly-correlated dataset} \label{weakly-mean}
    \begin{tabu}{|c|[1pt]c|c|c|c|c|c|}
        \tabucline[1pt]{1-6} 
        Testcase & 1 & 2 & 3 & 4 & 5 \\ [-1pt] \tabucline[1pt]{1-6} 
        DGHS Mean &1514$\pm$0 &1627.66$\pm$2.42 &8907.33$\pm$27.56 &17371.83$\pm$126.76 &82681.33$\pm$390.98 \\ \hline
        Mean Percent deviation from optimal &0\% &0.38\% &1.59\% &3.76\% &8.33\% \\[-1pt] \tabucline[1pt]{1-6}
        BMQHOA Mean &1508.16$\pm$6.82 &1613.66$\pm$16.74 &8435.33$\pm$60.60 &16539$\pm$133.95 &80900.66$\pm$970.71 \\ \hline
        Mean Percent deviation from optimal &0.38\% &1.24\% &6.81\% &8.37\% &10.31\% \\[-1pt] \tabucline[1pt]{1-6}
    \end{tabu}
\end{table}

\begin{figure}[hbt!]
    %\centering
    \begin{tikzpicture}
        \begin{axis} [
            title = Means of total profit values generated by DGHS and BMQHOA for the items weakly-correlated by profit-weight,
            ybar,
            bar width = 12pt,
            width = 14cm,
            height = 6cm,
            xlabel = Testcase,
            ylabel = Mean profit value,
            xmin = 1,
            xmax = 5,
            ymin=0,
            legend cell align={left},
            legend pos = north west,
            %x tick label style={rotate=45},
            xtick = data,
            %ytick = data,
            y tick label style={/pgf/number format/1000 sep=},
            enlarge y limits = {value = .25, upper},
            enlarge x limits = {abs = .8}
            %enlarge x limits = {abs = .8}
        ]
        \addplot+ [
                error bars/.cd,
                    y dir=both,
                    % (changed from `y explicit` so the error bars are (clearly) visible
                    y explicit ,
            ] coordinates {
                (1, 1514) +- (0, 0.00)
                (2, 1627.666667) +- (0, 2.422120283)
                (3, 8907.333333) +- (0, 27.56567914)
                (4, 17371.83333) +- (0, 126.7665834)
                (5, 82681.33333) +- (0, 390.9842282)
            };
        \addplot+ [
                error bars/.cd,
                    y dir=both,
                    % (changed from `y explicit` so the error bars are (clearly) visible
                    y explicit ,
            ] coordinates {
                (1, 1508.166667) +- (0, 6.823977335)
                (2, 1613.666667) +- (0, 16.74116683)
                (3, 8435.333333) +- (0, 60.6025302)
                (4, 16539) +- (0, 133.9596954)
                (5, 80900.66667) +- (0, 970.7196643)
            };
        \legend {DGHS, BMQHOA};
        \end{axis}
    \end{tikzpicture}
\caption{Shows the means of profit values for weakly correlated items} \label{weakly-graph}
\end{figure}

%strongly correlated
\begin{table}[h!]
    \centering
    \scriptsize
    \caption{\scriptsize Shows the mean and StDev. of the total profit values produced in the profit-weight strongly-correlated dataset} \label{strongly-mean}
    \begin{tabu}{|c|[1pt]c|c|c|c|c|c|}
        \tabucline[1pt]{1-6} 
        Testcase & 1 & 2 & 3 & 4 & 5 \\ [-1pt] \tabucline[1pt]{1-6} 
        DGHS Mean &2391.66$\pm$6.08 &2697$\pm$0 &13682.5$\pm$60.05 &26444.66$\pm$240.84 &126448$\pm$615.61 \\ \hline
        Mean Percent deviation from optimal &0.22\% &0\% &4.91\% &8.55\% &13.93\% \\[-1pt] \tabucline[1pt]{1-6}
        BMQHOA Mean &2327.33$\pm$48.55 &2658.66$\pm$48.81 &13334.66$\pm$182.91 &25604$\pm$91.32 &123813.83$\pm$773.19 \\ \hline
        Mean Percent deviation from optimal &2.90\% &1.42\% &7.33\% &11.46\% &15.72\% \\[-1pt] \tabucline[1pt]{1-6}
    \end{tabu}
\end{table}

\begin{figure}[hbt!]
    %\centering
    \begin{tikzpicture}
        \begin{axis} [
            title = Means of total profit values generated by DGHS and BMQHOA for the items strongly-correlated by profit-weight,
            ybar,
            bar width = 12pt,
            xlabel = Testcase,
            ylabel = Mean profit value,
            width = 14cm,
            height = 6cm,
            xmin = 1,
            xmax = 5,
            ymin = 0,
            legend cell align={left},
            legend pos = north west,
            %x tick label style={rotate=45},
            xtick = data,
            %ytick = data,
            y tick label style={/pgf/number format/1000 sep=},
            enlarge y limits = {value = .25, upper},
            enlarge x limits = {abs = .8}
            %enlarge x limits = {abs = .8}
        ]
        \addplot+ [
                error bars/.cd,
                    y dir=both,
                    % (changed from `y explicit` so the error bars are (clearly) visible
                    y explicit ,
            ] coordinates {
                (1, 2391.666667) +- (0, 6.08824003)
                (2, 2697) +- (0, 0)
                (3, 13682.5) +- (0, 60.05580738)
                (4, 26444.66667) +- (0, 240.8407496)
                (5, 126448) +- (0, 615.61384)
            };
        \addplot+ [
                error bars/.cd,
                    y dir=both,
                    % (changed from `y explicit` so the error bars are (clearly) visible
                    y explicit ,
            ] coordinates {
                (1, 2327.333333) +- (0, 48.55375028)
                (2, 2658.666667) +- (0, 48.81666382)
                (3, 13334.66667) +- (0, 182.9160099)
                (4, 25604) +- (0, 91.32579044)
                (5, 123813.8333) +- (0, 773.1955553)
            };
        \legend {DGHS, BMQHOA};
        \end{axis}
    \end{tikzpicture}
\caption{Shows the means of profit values for strongly correlated items} \label{strongly-graph}
\end{figure}

%mean percentage deviation
\begin{table}[h!]
    \centering
    \scriptsize
    \caption{\scriptsize Shows the mean percentage deviation from optimal total profit value across all datasets}\label{deviation-table}
    \begin{tabu}{|c|c|c|}
        \tabucline[1pt]{1-6} 
        & DGHS mean percentage deviation & BMQHOA mean percentage deviation \\ [-1pt] \tabucline[1pt]{1-6} 
        100 randomized items &0.12\% &4.17\% \\ \hline
        1000 randomized items &12.73\%&18.50\% \\\hline
        10k randomized items &22.78\%&25.01\% \\\hline
        Uncorrelated dataset &7.44\%&11.56\% \\\hline
        Weakly Correlated dataset &2.81\%&5.42\% \\\hline
        Strongly Correlated Dataset &5.52\%&7.77\% \\[-1pt] \tabucline[1pt]{1-6} 
    \end{tabu}
\end{table}

@@

        \legend {DGHS, BMQHOA};
        \end{axis}
    \end{tikzpicture}
    \caption{Mean percentage deviation from optimal across all datasets} \label{deviation-graph}
\end{figure}

\end{changemargin}

\clearpage
\subsection{Statistical Test}
An unpaired t-test was performed on the 2 groups of all the profit values produced by the DGHS and BMQHOA algorithms respectively to compare the means. The raw data was processed to get the means $\textcolor{purple}{x}$ of all profit values for each algorithm and the sum of squares $\textcolor{purple}{s}$ of the differences between observed values and the means for the profit values of each algorithm:

\begin{table}[h!]
    \centering
    \caption{\scriptsize Shows the means of the 2 groups of the profit values produced by the DGHS and BMQHOA algorithms, and the sums of squares of the differences between the observed values and the means of the groups} \label{t-test}
    \begin{tabu}{|c|[1pt]c|c|}
        \tabucline[1pt]{1-3} 
        & DGHS & BMQHOA \\ [-1pt] \tabucline[1pt]{1-6} 
        $\textcolor{purple}{x}$ & 1558988.34  &1513027.61 \\ \hline
        $\textcolor{purple}{s}$ & 2049263936280287.25 & 1934581405094429.75 \\\tabucline[1pt]{1-6}
    \end{tabu}
\end{table}

The formula for an unpaired t-test is:
$$ t = \frac{ x_{1} - x_{2} }{ \sqrt{ \frac{s_{1}^{2}}{n_{1}} + \frac{s_{2}^{2}}{n_{2}} } } $$
In this case, $n_{1} = n_{2} = 180$ as there are 180 values in both groups. This gives a t-test value of $0.00000000022$ and the corresponding p-value of $0.999998$. The null hypothesis is:
\begin{center}
    $H_{0}$ - There is no difference in total profit values produced by the DGHS and BMQHOA algorithms for the dataset.
\end{center}
